---
title: "330-HW-6"
author: "Chris Trippe"
date: "11/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, message = FALSE, echo = FALSE}
library(tidyverse)
library(MASS)
library(lmtest)
library(bestglm)
library(GGally)
library(car)
library(ggcorrplot)
library(kableExtra)
```

1. In your own words, summarize the overarching problem and any specific questions that need to be answered
using the rating dataset. Discuss how statistical modeling will be able to answer the posed questions.

In order to know what makes a good professor, we will make a regression model by examining nine different variables (number of years the professor has been teaching, number of courses taught by professor, is the professor "hot, the professor's discipline, average easiness rating of the professor, etc.) from ratemyprofessor.com and determine which, if any, contribute to a professor's overall quality. 

2. Construct a scatterplot matrix of the data and assess any potential issues related to collinearity. Calculate
variance inflation factors and discuss what variables, if any, are collinear. Comment on what affect collinearity
can have on a regression analysis.

```{r, message = FALSE, echo = FALSE}
##Read in table
prof_data <- read.table(file ="rateprof.txt", header = TRUE)

##Data subset with quantitative variables
prof_quant <- prof_data[,sapply(prof_data,is.numeric)]

##scatter plot matrix with correlations
ggpairs(prof_quant)
```


```{r, echo = FALSE}
##Linear Model with all variables
slr_prof <- lm(quality~.,prof_data)
vif_prof <- vif(slr_prof)
vif_prof <- data.frame(vif(slr_prof))
colnames(vif_prof) <- "VIF"
vif_prof <- as.matrix(cbind(Varibles = row.names(vif_prof), VIF = vif_prof$VIF))

##Calculate variance inflation factors
kable(cbind(vif_prof[1:10,], vif_prof[11:20,], vif_prof[21:30,])) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

When referencing to the scatterplot matrix and associated correlation values, there doesn't appear to be any collinearity or strong relationship between two explanitory variables. The largest correlation is 0.342 between number of years and number of raters, but we'll consider than relationship insignificant enough. When calculating variance inflation factors (VIF), which works for qualitative and quantitative variables, values above ten are considered to be multicollinear. That means that variables that should be significant will not be considered significant because of their relationship with other explanitory variables. For this data set, disciplinePre, disciplineSocSci, disciplineSTEM, deptBiology, deptBusiness, and deptChemistry all have VIFs above ten. 

```{r, echo = FALSE, results = "hide"}
##Calculate variance inflation factors > 10
vif(slr_prof)[vif(slr_prof) > 10]
```


3. Use a variable selection technique to determine a MLR model that will answer the questions posed in #1 (do
NOT consider any interactions). Justify your choice of your selected variable selection procedure (e.g., state
why you chose to use backward instead of forward selection). Justify your choice of a model comparison
criterion (e.g. state why you chose to base your variable selection procedure on AIC vs. BIC).

```{r}
prof_data <- prof_data[,c(1:7,9:ncol(prof_data),8)]

vs.res <- bestglm(prof_data, IC = "AIC", method = "exhaustive")

slr_best <- vs.res$BestModel
```

We will be using this model to make predictions which means AIC or Akaike Information Criterion, which is less restrictive on the variables allowed in "best" model, is better for making predictions than BIC or Bayesian Informaiton Criterion. The "exhaustive" selection method was also used because it finds the best model for each length one through the total number of variables based on their AIC values, and then of those models, the best is picked. That means every possible model is tested and the one with the best AIC value is chosen.

4. Write out (in mathematical form with greek letters) your selected MLR model from #3. Clearly state any
assumptions you are using in your model. Provide an interpretation of each mathematical term (variable or
parameter) included in your model. Using the mathematical form, discuss how your model, after fitting it to
the data, will be able to answer the questions in this problem.

When using the 
$y_i = \beta_0 + \beta_1numYears_i + \beta_2numCourses_i + \beta_3I(pepperyes_i = yes) + \beta_4easiness_i + \beta_5raterInterest_i + \beta_6I(disciplineSTEMYes_i = yes) + \beta_7I(deptBusinessYes_i = yes) + \beta_8I(deptEnglishYes_i = yes) + \beta_9I(deptGeologyYes_i = yes) + \beta_{10}I(deptLanguagesYes_i = yes) + \beta_{11}I(deptPhysicsYes_i = yes) + \beta{12}I(deptPolySciYes = yes) + \beta_{13}I(deptPsychologyYes_i = yes)$



```{r}

  
```


5. Fit your model in #3 to the professor rating data and summarize the results by displaying the estimated coefficients
in a table with 95% confidence intervals for each parameter. Provide an example of how to interpret
1 interval for a quantitative explanatory variable and 1 interval for a categorical explanatory variable correctly
in the context of the problem.

| Variable                             | Symbol        | Value                         | Variable Type |
| ------------------------------------ |-------------- | ----------------------------- | --------------|
| `r names(slr_best$coefficients[1])`  | $\beta_{0}$   | `r slr_best$coefficients[1]`  | Intercept     |
| `r names(slr_best$coefficients[2])`  | $\beta_{1}$   | `r slr_best$coefficients[2]`  | Quantitative  |
| `r names(slr_best$coefficients[3])`  | $\beta_{2}$   | `r slr_best$coefficients[3]`  | Quantitative  |
| `r names(slr_best$coefficients[4])`  | $\beta_{3}$   | `r slr_best$coefficients[4]`  | Qualitative   |
| `r names(slr_best$coefficients[5])`  | $\beta_{4}$   | `r slr_best$coefficients[5]`  | Quantitative  |
| `r names(slr_best$coefficients[6])`  | $\beta_{5}$   | `r slr_best$coefficients[6]`  | Quantitative  |
| `r names(slr_best$coefficients[7])`  | $\beta_{6}$   | `r slr_best$coefficients[7]`  | Qualitative   |
| `r names(slr_best$coefficients[8])`  | $\beta_{7}$   | `r slr_best$coefficients[8]`  | Qualitative   |
| `r names(slr_best$coefficients[9])`  | $\beta_{8}$   | `r slr_best$coefficients[9]`  | Qualitative   |
| `r names(slr_best$coefficients[10])` | $\beta_{9}$   | `r slr_best$coefficients[10]` | Qualitative   |
| `r names(slr_best$coefficients[11])` | $\beta_{10}$  | `r slr_best$coefficients[11]` | Qualitative   |
| `r names(slr_best$coefficients[12])` | $\beta_{11}$  | `r slr_best$coefficients[12]` | Qualitative   |
| `r names(slr_best$coefficients[13])` | $\beta_{12}$  | `r slr_best$coefficients[13]` | Qualitative   |
| `r names(slr_best$coefficients[14])` | $\beta_{13}$  | `r slr_best$coefficients[14]` | Qualitative   |

```{r}
confint(slr_best)
```

6. Assess the fit of your model and justify your model assumptions using appropriate graphics or summary
statistics. Provide discussion of your assessment of the model fit and assumptions on the level of your target
audience (e.g. interpret your model R2).

```{r, echo = FALSE}
#testing for linearity
avPlot(slr_best, variable = "numYears")
avPlot(slr_best, variable = "numCourses")
avPlot(slr_best, variable = "easiness")
avPlot(slr_best, variable = "raterInterest")

##testing for normality
ks.test(stdres(slr_best),"pnorm")

##testing for equal variance
bptest(slr_best)

```


7. Provide a discussion to an incoming BYU professor about important qualities of a good professor (according
to rate my professor anway). As an example, use your model to predict (and show a corresponding 95% prediction
interval) Dr. Heaton’s quality rating score given: numYears= 5, numRaters= 54, numCourses= 2,
pepper=no (§ - at least my wife thinks I’m cute) , discipline=STEM, dept=Math, easiness= 2.5,
raterInterest= 4.2.

